Comparing R packages for calculation of persistent homology

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised mem text

Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
Vietoris-Rips complices show a rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
Interestingly, different factors determine memory usage in the two types of simplicial complex.

@@

Figure \ref{fig:mem2}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size. Noisier shapes like the box and annulus increase object size compared to more uniform shapes like the sphere. 
Figure \ref{fig:mem2}, panel 1 shows that Rips complex object size depends only on the point cloud size and feature dimensions.

@@

<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)

Ripser memory not measured here b/c R, but refer to Bauer Ripser memory measurements.
-->

>>>>

Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
Vietoris-Rips complices show rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
Interestingly, different factors determine memory usage in the two types of simplicial complex.
Whereas memory use in calculation of a Vietoris-Rips complex is dependent on point cloud size and feature dimension, memory use for an analogous calculation with an alpha complex is also dependent on the structure of the point cloud itself (Figure \ref{fig:mem2}).
Within the limited set of tested point cloud structures, the sphere requires the least memory.

@@

@@

<!--Include somewhere:

differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)

Ripser memory not measured here b/c R, but refer to Bauer Ripser memory measurements.

Rips complex memory more predictable, but alpha complex memory generally better (but what about Ripser?).

Thoughts on why sphere alpha uses less memory than 3-d annulus alpha
-->
